---
title: "Ames Housing Data Analysis"
author: "Jaykumar Patel"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
library(kableExtra)
```

## Load Libraries

```{r, include=FALSE}

knitr::opts_chunk$set(fig.width=10, fig.height=6, fig.align = "center")

```

```{r packages, warning = FALSE, message = FALSE}
library(scales)
library(tidyverse)
library(modelsummary)
library(ggthemes)
library(gridExtra)
library(broom)
```

```{r, include=FALSE}

themes <-  function()
  {
  theme(
        plot.title = element_text(color = "#FF9671", size = 12, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, color = "#845EC2",  face = "italic" ),
        axis.title.x = element_text(color = "#FF9671", face = "bold",margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.title.y = element_text(color = "#FF9671", face = "bold", margin = margin(t = 0, r = 10, b = 0, l = 0)),
        text = element_text(colour = "#BCF7E4"),
        plot.background = element_rect(fill = "#EDFFF9", color = NaN, margin(t = 10, r = 10, b = 10, l = 10)),
        panel.background = element_rect(fill = "#DDFFF4"),
        panel.grid.major = element_line(color = "#BCF7E4"),
        panel.grid.minor = element_line(color = "#BCF7E4"),
        legend.background = element_rect(fill="#B0E0E0",size=0.5, linetype="solid"),
        legend.title = element_text(colour="black", size=10, face="bold"),
        legend.text = element_text(color = "#545479"),
        legend.position = "none",
        legend.direction = "horizontal" 
        )
}

colors <- c('#845EC2', '#D65DB1', '#FF6F91', '#FF9671', "#FFC75F", "#A8F78B","#9367C4", "#4B4453")

```

## Load Ames Housing Data

```{r}

data(ames, package = "modeldata")

```

## Select Numeric Variables

```{r}
num <- ames %>%
  select(where(is.numeric))

head(num)
```


# EDA

## Exploring the univariate distribution of the response variable Sale price.

```{r}
ames %>% 
  ggplot(aes(Sale_Price))+    
  geom_histogram(aes(y = stat(density)), color = "#13B4FA",fill = "#FF6F91") +
    geom_density(fill = "#845EC2", alpha = 0.5, color = NaN)+
    geom_vline(xintercept = mean(ames$Sale_Price), lwd = 0.8, linetype = "dashed", color = "#845EC2" )+
     geom_label(label= paste("Mean", round(mean(ames$Sale_Price), 2), sep =": "), x=250000, y=0.000007, label.padding = unit(0.55, "lines"), label.size = 0.25, fill = "#845EC2", color = "white"
  )+
    geom_vline(xintercept = median(ames$Sale_Price), lwd = 0.8, linetype="twodash", color = "#4B4453")+
     geom_label(label= paste("Median", round(median(ames$Sale_Price), 2), sep =": "), x=100000, y=0.000006, label.padding = unit(0.55, "lines"), label.size = 0.25, fill = "#4B4453", color = "white"
  )+
    labs(title = "Distribution of Sales Price",
         x = "Total Price",
         y = "density") +
  scale_x_continuous(labels  = label_number(scale = 1e-05, prefix = "$", suffix = "hk", accuracy = 1))+
    theme_replace()+
    themes()
```
                  
-   As we look at the mean and median, `median < mean` so we can say that our variable is right skewed. But it is not much skewed so we can consider it as normally distributed because in real world we never get perfectly distributed data.        
-   The reason behind this skewness can be outliers present in the response variable..


## Find outlier(s) in your original response variable

```{r}
ames %>% 
  ggplot(aes(y= Sale_Price))+
  geom_boxplot(outlier.color = "#13B4FA" )+
  labs(title = "Outliers in Sale Price",
       y = "Sale Price") +
  theme_replace()+
  scale_y_continuous(labels  = label_number(scale = 1e-05, prefix = "$", suffix = "hk", accuracy = 1))+
  themes()

```

- We have found the we have some outliers present in the data which can be reason for the skewness.

# Bivariate analysis

## Relation between Sale Price and Total Basement size

```{r}

data1 <- ames %>% 
          filter((!Total_Bsmt_SF>4000)& (Total_Bsmt_SF > 0)) %>% 
          select(Total_Bsmt_SF,Sale_Price)

bsmt1 <- data1 %>% 
  ggplot(aes(Total_Bsmt_SF, Sale_Price)) +
  geom_point() +
  labs(title = "Relation between Sale Price and Total Basement size(with outliers)",
         x = "Total Basement size",
         y = "Sale Price")+
  theme_replace() +
      scale_y_continuous(labels  = label_number(scale = 1e-05, prefix = "$", suffix = "hk", accuracy = 1)) +
  themes()

bsmt2 <- ames %>% 
  ggplot(aes(Total_Bsmt_SF, Sale_Price)) +
  geom_point() +
  labs(title = "Relation between Sale Price and Total Basement size(without outliers)",
         x = "Total Basement size",
         y = "Sale Price")+
  theme_replace() +
      scale_y_continuous(labels  = label_number(scale = 1e-05, prefix = "$", suffix = "hk", accuracy = 1)) +
  themes()

grid.arrange(bsmt1, bsmt2, ncol=2)

```

```{r}

Bsmt_after = cor(data1['Total_Bsmt_SF'], data1['Sale_Price'])

Bsmt_before = cor(ames['Total_Bsmt_SF'], ames['Sale_Price'])

sprintf("Correlation coeffeicient before removing outliers: %f", Bsmt_before)
sprintf("Correlation coeffeicient After removing outliers: %f", Bsmt_after)

```

- Here, we can say that we have some positive relation between sale price and total basement size.

- It's correlation coefficient is : `r cor(data1['Total_Bsmt_SF'], data1['Sale_Price'])`, which is showing moderate relation.


























